---
title: "Matching and Subclassification"
author: "akash.yadav"
---
# Load packages
```{r}
#| warning: false
library(tidyverse)
library(dagitty)
library(ggdag)
library(MatchIt)
```
# Load data
```{r}
data<-readRDS(" ../../../Causal_Data_Science_Data/membership.rds")
```


# Assignment 1
## Plotting the data to check the relationships
```{r}
pla <- ggplot(data[,c(1,5)], aes(x = age, y = avg_purch)) +
  geom_point(alpha = .8) +
  stat_smooth(method = "lm", se = F)

pla

ppp<- ggplot(data[,c(1,3)], aes(x = age, y = pre_avg_purch)) +
  geom_point(alpha = .8) +
  stat_smooth(method = "lm", se = F)

pla_card<- ggplot(data[,c(1,4)], aes(x = age, y = card)) +
  geom_point(alpha = .8) +
  stat_smooth(method = "lm", se = F)

ppap_card<- ggplot(data[,c(3,4)], aes(x = pre_avg_purch, y = card)) +
  geom_point(alpha = .8) +
  stat_smooth(method = "lm", se = F)

ppp
pla_card
ppap_card
```

## DAG
```{r}
# Define DAG structure
dag_model <- 'dag {
bb="0,0,1,1"
age [exposure,pos="0.075,0.4", color = "skyblue"]
avg_purch [outcome,pos="0.4,0.4", color = "orange"]
card [pos="0.2,0.2", color = "purple"]
pre_avg_purch [pos="0.2,0.6", color = "lightgreen"]
age -> avg_purch
age -> pre_avg_purch
pre_avg_purch -> card
age -> card
card -> avg_purch
pre_avg_purch -> avg_purch
}
'

# Create DAG object
dag <- dagitty(dag_model)

# Plot DAG with ggraph
ggdag(dag) +
  geom_dag_point() +
  geom_dag_text() +
  geom_dag_edges() +
  theme_dag_blank() +
  geom_dag_text(colour = "orange")
```


# Assignment 2
```{r}
lm_2<-lm(avg_purch~card,data)
summary(lm_2)
```
The naive estimate of the treatment effect is about 25.22.



# Assignment 3.1
```{r}
cem <- matchit(card ~ age + pre_avg_purch, # Coarsened Exact Matching
               data = data, 
               method = 'cem', 
               estimand = 'ATE')
summary(cem) # Covariate balance

df_cem <- match.data(cem) # Use matched data
model_cem <- lm(avg_purch ~ card, data = df_cem, weights = weights) # Estimation
summary(model_cem)
```
After the Coarsened Exact Matching the estimate for the ATE is 15.27.



# Assignment 3.2
```{r}
nn <- matchit(card ~ age + pre_avg_purch, # Nearest-Neighbor Matching
              data = data,
              method = "nearest",
              distance = "mahalanobis",
              replace = T)
summary(nn) # Covariate Balance

df_nn <- match.data(nn) # Use matched data
model_nn <- lm(avg_purch ~ card, data = df_nn, weights = weights) # Estimation
summary(model_nn)
```
After the Nearest-Neighbor Matching the estimate for the ATE is 14.95.



# Assignment 3.3
```{r}
# Code for propensity scores and IPW modeling
model_prop <- glm(card ~ age + pre_avg_purch, # (1) Propensity scores
                   data = data,
                   family = binomial(link = "logit"))
summary(model_prop)

df_aug <- data %>% mutate(propensity = predict(model_prop, type = "response")) # Add propensities to table
df_ipw <- df_aug %>% mutate(ipw = (card/propensity) + ((1-card) / (1-propensity))) # Extend data by IPW scores
df_ipw %>% select(card, age, pre_avg_purch, propensity, ipw) # Look at data with IPW scores

model_ipw <- lm(avg_purch ~ card, # (2) Estimation
                  data = df_ipw,
                  weights = ipw)
summary(model_ipw)

# Plot propensity scores as a density plot
ggplot(df_aug, aes(x = propensity)) +
  geom_density(alpha = 0.8)

# Plot propensity scores as a boxplot
ggplot(df_aug, aes(x = card, y = propensity)) +
  geom_boxplot()

# Plot IPW scores as a scatterplot
ggplot(df_ipw, aes(x = card, y = avg_purch, color = card)) +
  geom_point()

# Plot relationship between average purchase and card using IPW weights
ggplot(df_ipw, aes(x = card, y = avg_purch, color = card)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

## Summary of findings

Analysis revealed a positive association between credit card ownership and average purchase amount. After adjusting for potential confounding factors using inverse probability weighting, found that individuals with credit cards tend to spend an average of **€14.95** more per purchase than those without cards.

## Key takeaways

* Credit card usage is linked to higher average spending.

* The IPW-adjusted effect size suggests a measurable impact of credit cards on spending habits.

* While credit cards may influence spending, other factors also play a role.

