---
title: "Regression Discontinuity"
author: "akash.yadav"
---

> Loading the data frame

```{r}
#| warning: false
library(dplyr)

dataFrameCoupoun <- readRDS("../Causal_Data_Science_Data/coupon.rds")

dataFrameCoupoun
```

# Assignment - 1

------------------------------------------------------------------------

:::{.callout-important}
with half the bandwidth
:::

```{r}
# Define cut-off
c0 <- 60
# Specify bandwidth
bw <- (c0 + c(-2.5, 2.5))

# Subsets below and above threshold in specified bandwidth
DWBelow <- dataFrameCoupoun %>% filter(days_since_last >= bw[1] & days_since_last < c0)
DWAbove <- dataFrameCoupoun %>% filter(days_since_last >= c0 & days_since_last <= bw[2])

DataFrameBW <- bind_rows(DWAbove, DWBelow)
dim(DataFrameBW)
```

> LATE

```{r}
model_bw_below <- lm(purchase_after ~ days_since_last, DWBelow)
model_bw_above <- lm(purchase_after ~ days_since_last, DWAbove)

y0 <- predict(model_bw_below, tibble(days_since_last = c0))
y1 <- predict(model_bw_above, tibble(days_since_last = c0))

late <- y1 - y0
sprintf("LATE: %.2f", late)
```

> Ploting 

```{r}
library(ggplot2)

# Minimum and maximum for y-axis limits
min_y <- min(DataFrameBW$purchase_after)
max_y <- max(DataFrameBW$purchase_after)

# Add lines for vertical distance and change limits of x-axis.
dep_var_bw <- 
  ggplot(DataFrameBW, aes(x = days_since_last, y = purchase_after, color = coupon)) +
  geom_vline(xintercept = c0, linewidth = 2) +
  geom_point(alpha = 0.4, size = 1) +
  geom_smooth(data = DWBelow, method = "lm", se = F, linewidth = 2) +
  geom_smooth(data = DWAbove, method = "lm", se = F, linewidth = 2) +
  geom_segment(aes(x = c0, xend = bw[2], y = y0, yend = y0),
             linetype = "dotted") +
  geom_segment(aes(x = bw[1], xend = c0, y = y1, yend = y1),
               linetype = "dotted") +
  annotate("text", x = c0+2, y = mean(c(y1, y0)-2),
           label = sprintf("Difference: %.2f", (y1 - y0)),
           fontface = 2) +
  scale_y_continuous(limits = c(min_y, max_y)) + 
  scale_color_discrete(labels = c("No coupon", "Coupon")) +
  xlab("Days since last purchase") +
  ylab("Purchase after coupon assignment") +
  theme(legend.title = element_blank())
dep_var_bw
```


```{r}
# Specify bandwidth
bw <- (c0 + c(-10, 10))

# Subsets below and above threshold in specified bandwidth
DWBelow <- dataFrameCoupoun %>% filter(days_since_last >= bw[1] & days_since_last < c0)
DWAbove <- dataFrameCoupoun %>% filter(days_since_last >= c0 & days_since_last <= bw[2])

DataFrameBW <- bind_rows(DWAbove, DWBelow)
dim(DataFrameBW)
```

> LATE

```{r}
model_bw_below <- lm(purchase_after ~ days_since_last, DWBelow)
model_bw_above <- lm(purchase_after ~ days_since_last, DWAbove)

y0 <- predict(model_bw_below, tibble(days_since_last = c0))
y1 <- predict(model_bw_above, tibble(days_since_last = c0))

late <- y1 - y0
sprintf("LATE: %.2f", late)
```

> Ploting

```{r}
# Minimum and maximum for y-axis limits
min_y <- min(DataFrameBW$purchase_after)
max_y <- max(DataFrameBW$purchase_after)

# Add lines for vertical distance and change limits of x-axis.
dep_var_bw <- 
  ggplot(DataFrameBW, aes(x = days_since_last, y = purchase_after, color = coupon)) +
  geom_vline(xintercept = c0, linewidth = 2) +
  geom_point(alpha = 0.4, size = 1) +
  geom_smooth(data = DWBelow, method = "lm", se = F, linewidth = 2) +
  geom_smooth(data = DWAbove, method = "lm", se = F, linewidth = 2) +
  geom_segment(aes(x = c0, xend = bw[2], y = y0, yend = y0),
             linetype = "dotted") +
  geom_segment(aes(x = bw[1], xend = c0, y = y1, yend = y1),
               linetype = "dotted") +
  annotate("text", x = c0+2, y = mean(c(y1, y0)-2),
           label = sprintf("Difference: %.2f", (y1 - y0)),
           fontface = 2) +
  scale_y_continuous(limits = c(min_y, max_y)) + 
  scale_color_discrete(labels = c("No coupon", "Coupon")) +
  xlab("Days since last purchase") +
  ylab("Purchase after coupon assignment") +
  theme(legend.title = element_blank())
dep_var_bw
```

**Bandwidth Sensitivity and LATE Values**

The choice of bandwidth exerts a significant influence on the estimated LATE value. Increasing the bandwidth leads to a higher LATE estimate, while decreasing it results in a slightly lower one. A broader bandwidth encompasses a wider range of data, potentially introducing more variation, whereas a smaller bandwidth might fail to capture sufficient treatment effects, leading to a lower estimate. This sensitivity to bandwidth underscores its crucial role in determining LATE values.


# Assignment - 2

:::{.callout-important}
with double the bandwidth
:::


------------------------------------------------------------------------

> Loading the data frame

```{r}
df_shipping  <- readRDS("../Causal_Data_Science_Data/shipping.rds")
df_shipping
```

> Plot

```{r}
df_shipping <- df_shipping %>%
  mutate(shipping_type = ifelse(purchase_amount > 30, "Zero-cost Shipping", "cost Shipping"))

ggplot(df_shipping, aes(x = purchase_amount, y = shipping_type, color = shipping_type)) +
  geom_vline(xintercept = 30, color = palette()[2], linetype = "dashed") +
  geom_point(alpha = 0.2, position = position_jitter()) +
  guides(scale = "none") +
  scale_y_discrete(labels = c("Zero-cost Shipping", "cost Shipping"))+
  scale_color_discrete(labels = c("Zero-cost Shipping", "cost Shipping")) +
  xlab("Purchase Amount") +
  ylab("") +
  theme(legend.position = "none")
```

**Supporting the Claim with Data Visualization**

The effectiveness of the argument is corroborated by the distinct 30€ cutoff line and the spread of purchases on either side. This aligns with the provided information on free shipping for purchases above 30€. The dotted line at 30€ aptly marks the cutoff point. The chart effectively delineates how purchase amount differentiates between purchases qualifying for free shipping and those bearing shipping costs.

In summary, the data visualization provides strong evidence to support the claim that purchase amount serves as a reliable indicator of shipping type, with purchases exceeding 30€ generally eligible for free shipping and those below 30€ typically incurring shipping costs. The clear demarcation of the 30€ cutoff and the scattered distribution of purchases on either side further reinforce this conclusion.
